### 2016-01-08 first_response_to_referee.txt
Started to write the response to the referee.

Bella Lock (Mrs) 
Assistant Editor, Royal Astronomical Society

Dear Mrs. Bella,

 I am resubmitting the manuscript MNRAS/MN-15-3639-MJ:

 "Limb-darkening and exoplanets II: Choosing the Best 
  Law for Minimum Bias Retrieval of Transit Parameters"

 Authors: Nestor Espinoza & Andres Jordan

 We thank the referee for his/her careful reading of the manuscript. 
 You will find below our detailed response to the referee's comments 
 as indented text after each of his/her comments.

 Besides the changes made in response to the referee's comments we have:

     - Corrected minor typos on the manuscript.

########################################################################################
>> My main suggestion is that the paper would benefit from a more quantitative 
>> description of the trade-off between precision and accuracy. This work indicates 
>> that fitting for two limb darkening coefficients with a judiciously chosen model 
>> can significantly reduce bias. In practice, however, many light curves are not 
>> sufficiently precise to tightly constrain more than one limb darkening parameter. 
>> Adding an additional degree of freedom with a second limb darkening coefficient 
>> may cause the uncertainty on other important parameters, particularly the transit 
>> depth, to blow up. The resulting loss in accuracy may not be worth the improvement 
>> in precision due to choosing a two-parameter model over fitting for  a linear limb 
>> darkening coefficient.
>>
>> One way to address this question is to give guidelines about how precise a 
>> light curve must be to merit adding a second limb darkening parameter to the 
>> model. This could be discussed for a few select examples (e.g., a hot Jupiter 
>> transiting a Sun-like star).  I leave it up to the authors whether/how to address 
>> this point.

RESPONSE

>> It would also be helpful to investigate how much bias is introduced by using a linear 
>> limb darkening law, since this model is commonly used in practice. Again, the authors 
>> can address this at their discretion.

RESPONSE

>> I would recommend adding a few sentences to the abstract to give a little more context 
>> for the paper. For example, something to the effect of: "Very precise measurements of 
>> exoplanet transit light curves enable limb darkening parameters to be fitted empirically, 
>> rather than fixed on theoretical model predictions that may have unknown systematic errors. 
>> However, even fitting for the limb darkening coefficients can introduce bias in the measured 
>> transit parameters depending on the chosen limb darkening model parameterization." I recommend 
>> adding this because because fitting for the coefficients directly is still a relatively new 
>> approach in this field, and it may not be obvious to everyone why you're not just using the 
>> standard nonlinear law.

RESPONSE

>> There are several cases of very precise transit light curves where the limb darkening coefficients 
>> predicted from theoretical models does not provide a good fit to the data (Knutson et al. 2007, ApJ, 
>> 655, 564; Kreidberg et al. 2015, ApJ, 814, 66).  Please cite these in the introduction.

RESPONSE

>> State in the text what time sampling you use for your model light curves, and how sensitive your 
>> results are to the choice of time sampling.

RESPONSE

>> There are a few uneccessary lines in the code run_ld_exosim.py that should be removed:
>>      import pymc
>>      sys.path.append('../utilities/')

RESPONSE