### 2016-01-15 first_response_to_referee.txt
Finished response to the referee. Added two new simulations: one with the linear law, and one on the bias/variance trade-off. Edits and comments are welcome.

Bella Lock (Mrs) 
Assistant Editor, Royal Astronomical Society

Dear Mrs. Bella,

 I am resubmitting the manuscript MNRAS/MN-15-3639-MJ:

 "Limb-darkening and exoplanets II: Choosing the Best 
  Law for Minimum Bias Retrieval of Transit Parameters"

 Authors: Nestor Espinoza & Andres Jordan

 We thank the referee for his/her careful reading of the manuscript. 
 You will find below our detailed response to the referee's comments 
 as indented text after each of his/her comments.

 Besides the changes made in response to the referee's comments we have:

     - Corrected minor typos on the manuscript.

     - Changed the title of the manuscript to reflect the additional study 
       proposed by the referee on the precision of the retrieved transit 
       parameters, which implies we now not only help to decide the law 
       that provide minimum bias, but also the one that provides the 
       best precision. This in turn means that we help to decide the 
       optimal law, which is a statistical jargon for "lowest bias, 
       lowest variance" model.

     - Added extra text on the Abstract, Introduction and Section 3 to reflect 
       these changes in the study. This includes a change of the name of Section 3, 
       and an additional sub-section explaining the procedures used to study the 
       bias/variance trade-off.

     - Modified the Acknowledgments.

########################################################################################
>> My main suggestion is that the paper would benefit from a more quantitative 
>> description of the trade-off between precision and accuracy. This work indicates 
>> that fitting for two limb darkening coefficients with a judiciously chosen model 
>> can significantly reduce bias. In practice, however, many light curves are not 
>> sufficiently precise to tightly constrain more than one limb darkening parameter. 
>> Adding an additional degree of freedom with a second limb darkening coefficient 
>> may cause the uncertainty on other important parameters, particularly the transit 
>> depth, to blow up. The resulting loss in accuracy may not be worth the improvement 
>> in precision due to choosing a two-parameter model over fitting for  a linear limb 
>> darkening coefficient.
>>
>> One way to address this question is to give guidelines about how precise a 
>> light curve must be to merit adding a second limb darkening parameter to the 
>> model. This could be discussed for a few select examples (e.g., a hot Jupiter 
>> transiting a Sun-like star).  I leave it up to the authors whether/how to address 
>> this point.

We thank the referee for this very insightful suggestion. We have made a whole suit of 
simulations which exactly answers this question, and provide code so future researchers 
can decide which is the best law both in terms of bias and precision for their specific 
application; in this way, researchers can decide the optimal limb-darkening law for the 
lightcurve at hand. These simulations are shown and discussed both on Section 3.4 and 
Section 4.

>> It would also be helpful to investigate how much bias is introduced by using a linear 
>> limb darkening law, since this model is commonly used in practice. Again, the authors 
>> can address this at their discretion.

We thank the referee for making this suggestion. We have added simulations that provide 
the results for the linear law. As expected, the bias is larger than for any of the other 
laws, and this is now discussed both in the abstract and in Sections 3 and 4.

>> I would recommend adding a few sentences to the abstract to give a little more context 
>> for the paper. For example, something to the effect of: "Very precise measurements of 
>> exoplanet transit light curves enable limb darkening parameters to be fitted empirically, 
>> rather than fixed on theoretical model predictions that may have unknown systematic errors. 
>> However, even fitting for the limb darkening coefficients can introduce bias in the measured 
>> transit parameters depending on the chosen limb darkening model parameterization." I recommend 
>> adding this because because fitting for the coefficients directly is still a relatively new 
>> approach in this field, and it may not be obvious to everyone why you're not just using the 
>> standard nonlinear law.

We thank the referee for this suggestion. We have rewritten the first few sentences in the abstract 
in order to provide the suggested context to the paper.

>> There are several cases of very precise transit light curves where the limb darkening coefficients 
>> predicted from theoretical models does not provide a good fit to the data (Knutson et al. 2007, ApJ, 
>> 655, 564; Kreidberg et al. 2015, ApJ, 814, 66).  Please cite these in the introduction.

We have added the suggested citations to the discussion in the introduction.

>> State in the text what time sampling you use for your model light curves, and how sensitive your 
>> results are to the choice of time sampling.

We have added the following at the end of the second paragraph in Section 3:

"$1000$ in-transit points and $400$ out-of-transit points were simulated per 
lightcurve, with the initial times perturbed by a small amount in order for 
the results to be independent of the exact points at which the lightcurve is sampled 
(which anyways our simulations show has a negligible effect given the large amount of 
sampled points)."

>> There are a few uneccessary lines in the code run_ld_exosim.py that should be removed:
>>      import pymc
>>      sys.path.append('../utilities/')

We thank the referee for catching this typo in the code. The lines have been removed from it.